---
title: "SQL Review"
author: "Cari Kaufman"
date: "February 9, 2015"
output: html_document
---

I'm assuming you are already familiar with the material from the SQL Bootcamp at  <http://software-carpentry.org/v5/novice/sql/>. If not, you'll need to take some time to go through it on your own. We'll review the syntax of the SELECT comand, with particular emphasis on

1.  Constructing queries in the order they'll be evaluated
2.  Using subqueries
3.  Joining more than one table.

Because I'll be using the results of the queries to make plots in R, I'm going to connect to the database from within R, using the package RMySQL. (The username and password will be given in class.)

```{r, echo=FALSE}
user = "****"; password = "****" # Replace with info given in lecture
```
```{r}
library(RMySQL)
con = dbConnect(RMySQL::MySQL(), dbname = "ontime",
                host = "radagast.berkeley.edu",
                user = user, password = password)
```

Let's see what tables exist in the database.

```{r}
dbListTables(con)
```
We can look at the columns/attributes in a particular table.
```{r}
dbListFields(con, "ontime")
```
## Writing queries in order of evaluation

Let's first review SELECT commands that work with a single table. The basic format looks like this.

```
SELECT columns or aggregate functions of columns
FROM table
[WHERE conditions]
[GROUP BY columns]
[HAVING conditions]
[ORDER BY columns [ASC | DESC] ]
[LIMIT rows]
```

The thing that has helped me the most in learning to write SQL queries is to write them in the order in which they're going to be evaluated. Breaking them down in this way gives you a series of simple questions to answer. The order in which the SELECT command above is evaluated is

1. FROM: The working table is constructed.
2. WHERE: The WHERE clause is applied to each row of the table, and only the rows that test TRUE are retained.
3. GROUP BY: The results are broken into groups of rows all with the same value of the GROUP BY clause.
4. HAVING: The HAVING clause is applied to each group and only those that test TRUE are retained.
5. SELECT: The attributes not in the list are dropped, aggregates are calculated, and options DISTINCT, ORDER BY, and LIMIT are applied.

### Example 1

*Working with the airports table, calculate the number of airports in each state. Report the results in decreasing order of the number of airports.*

First a quick look at what's in the airports table.
```{r}
dbListFields(con, "airports")
```

Let's fill of the 5 steps in turn. I'll show you what my query looks like after each step.

1. FROM: We need data only from the airports table.
```
FROM airports
```
2. WHERE: The question needs all the rows, so there's nothing to add here.
3. GROUP BY: We want to create a subtable for each state and then count the number of rows in it.
```
FROM airports
GROUP BY state
```
4. HAVING: The question includes all states, so there's nothing to add here.
5. SELECT: We want a table with a column for state and one for the total number of airports, which is the same as the number of rows. Note that we can rename a column using AS. We also need to ORDER BY this new column.
```
SELECT state, COUNT(*) AS total
FROM airports
GROUP BY state
ORDER BY total DESC;
```

Ok, let's run it in R!

```{r}
query = 
"SELECT state, COUNT(*) AS total
FROM airports
GROUP BY state
ORDER BY total DESC;"
statetotals = dbGetQuery(con, query)
statetotals
```
It's interesting that Alaska has the most airports. I would guess that many of them are very small. Thinking ahead, we might also want to look at the total volume of flights into and out of each airport.

Clearly there are also some values in the state column other than the 50 states.
```{r}
setdiff(statetotals$state, state.abb)
```
### Example 2

*Calculate the average departure delay out of SFO in 2008 for each day of the week. Only include the delayed flights in the average (i.e., DepDelay > 0).*

1. FROM: Just need the ontime table.
```
FROM ontime
```
2. WHERE: Remove flights that aren't delayed. Only include 2008 and SFO.
```
FROM ontime
WHERE DepDelay > 0 AND Year = 2008 AND Origin = SFO
```
3. GROUP BY: Make subtables for day of week.
```
FROM ontime
WHERE DepDelay > 0 AND Year = 2008 AND Origin = SFO
GROUP BY DayOfWeek
```
4. HAVING: I don't want to exclude any day of week, so there's nothing to do here. 
5. SELECT: 
```
SELECT DayOfWeek, AVG(DepDelay) AS AvgDelay
FROM ontime
WHERE DepDelay > 0 AND Year = 2008 AND Origin = SFO
GROUP BY DayOfWeek
ORDER BY DayOfWeek;
```
Let's run it in R.

```{r}
query = 
"SELECT DayOfWeek, AVG(DepDelay) AS AvgDelay
FROM ontime
WHERE DepDelay > 0 AND Year = 2008 AND Origin = 'SFO'
GROUP BY DayOfWeek
ORDER BY DayOfWeek;"
SFODayOfWeek2008 = dbGetQuery(con, query)
SFODayOfWeek2008
```
### Example 3

*Find the worst 5 airports in terms of departure delays in 2008 and their average delays. This time include all flights (early, on time, and delayed) in the average.*

1. FROM: We just need the ontime table.
```
FROM ontime
```
2. WHERE: Just keep rows for 2008.
```
FROM ontime
WHERE Year = 2008
```
3. GROUP BY: Make subtables for each airport (Origin)
```
FROM ontime
WHERE Year = 2008
GROUP BY Origin
```
4. HAVING: Nothing to do here.
5. SELECT: 
```
SELECT Origin, AVG(DelDelay) AS AvgDelay
FROM ontime
WHERE Year = 2008
GROUP BY Origin
ORDER BY AvgDelay DESC
LIMIT 5;
```

```{r}
query = 
"SELECT Origin, AVG(DepDelay) AS AvgDelay
FROM ontime
WHERE Year = 2008
GROUP BY Origin
ORDER BY AvgDelay DESC
LIMIT 5;"
Worst5 = dbGetQuery(con, query)
Worst5
```

## Using subqueries

A subquery is just a query whose output is used as part of a more complicated query. It's very common for these to show up in the WHERE or HAVING clauses. Here's an example to illustrate this.

### Example 4

*Find how many flights left SFO in 2008. Then make a list of all the airports with more than this many flights in 2008.*

We can build up the first query as follows.

1. FROM: We just need the ontime table.
```
FROM ontime
```
2. WHERE: Restrict to 2008 and SFO.
```
FROM ontime 
WHERE Year = 2008 and Origin = 'SFO'
```
3. GROUP BY: Nothing to do here.
4. HAVING: Ditto.
5. SELECT: Count the number of rows (flights).
```
SELECT COUNT(*)
FROM ontime 
WHERE Year = 2008 and Origin = 'SFO';
```

```{r}
query = 
"SELECT COUNT(*)
FROM ontime 
WHERE Year = 2008 and Origin = 'SFO';"
dbGetQuery(con, query)
```

Moving onto the second query, we have

1. FROM: Still working with the ontime table.
```
FROM ontime
```
2. WHERE: Restrict to 2008.
```
FROM ontime
WHERE Year = 2008
```
3. GROUP BY: Split things up by origin.
```
FROM ontime
WHERE Year = 2008
GROUP BY Origin
```
4. HAVING: Keep only airports with more flights than SFO
```
FROM ontime
WHERE Year = 2008
GROUP BY Origin
HAVING COUNT(*) > (SELECT COUNT(*) FROM ontime 
                   WHERE Year = 2008 and Origin = 'SFO')
```
5. SELECT: The question asked for the airport names. I'll go ahead and include the number of flights, too.
```
SELECT Origin, COUNT(*) AS Flights
FROM ontime
WHERE Year = 2008
GROUP BY Origin
HAVING COUNT(*) > (SELECT COUNT(*) FROM ontime 
                   WHERE Year = 2008 and Origin = 'SFO');
```

```{r}
query = 
"SELECT Origin, COUNT(*) AS Flights
FROM ontime
WHERE Year = 2008
GROUP BY Origin
HAVING COUNT(*) > (SELECT COUNT(*) FROM ontime 
                   WHERE Year = 2008 and Origin = 'SFO');"
morethanSFO = dbGetQuery(con, query)
morethanSFO
```

## Joining tables

I believe you've already covered this. I want to show you a complex example that combines it with a subquery. 

### Example 5

*Repeat example 3, but this time include only airports with more than 50000 total flights, and also include all the location information from the airports table.*

We can accomplish this task by first doing everything except the location information first, as a subquery. Then join the result of that subquery with the airports table. Note that if you want to refer back to the columns of a subquery, you need to name it, i.e. (subquery) AS x and then x.column.

1. FROM: Data in the ontime table and airports
```
FROM (SELECT Origin, AVG(DepDelay) AS AvgDelay
      FROM ontime
      WHERE Year = 2008
      GROUP BY Origin
      HAVING COUNT(*) > 50000
      ORDER BY AvgDelay DESC
      LIMIT 5) AS x JOIN airports ON x.Origin = airports.iata
```
2. WHERE: Nothing to do here.
3. GROUP BY: Nothing to do here.
4. HAVING: Nothing to do here.
5. SELECT: Keep everything from x and airports
```
SELECT *
FROM (SELECT Origin, AVG(DepDelay) AS AvgDelay
      FROM ontime
      WHERE Year = 2008
      GROUP BY Origin
      HAVING COUNT(*) > 50000
      ORDER BY AvgDelay DESC
      LIMIT 5) AS x JOIN airports ON x.Origin = airports.iata;
```

```{r}
query = 
"SELECT *
FROM (SELECT Origin, AVG(DepDelay) AS AvgDelay
      FROM ontime
      WHERE Year = 2008
      GROUP BY Origin
      HAVING COUNT(*) > 50000
      ORDER BY AvgDelay DESC
      LIMIT 5) AS x JOIN airports ON x.Origin = airports.iata;"
Worst5Updated = dbGetQuery(con, query)
Worst5Updated
```
